---
title: "Lecture 10"
author: "Brandon Foltz"
date: "2023-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom) #tidy(), augment(), glance()
library(ggeffects)
library(pander)
library(gt)
library(modelsummary)
library(gtsummary)
library(corrplot)
mtcars
```
```{r}
lm_cars <- lm(mpg ~ disp + wt, data = mtcars)
summary(lm_cars)
```
 
```{r}
glm_cars <- glm(am ~ disp + wt, data = mtcars, family = binomial(link = "logit"))
summary(glm_cars)
```

```{r}
lm_cars %>% tidy()
glm_cars %>% tidy()
```

```{r}
glance(lm_cars)
glance(glm_cars)
```

```{r}
augment(lm_cars) %>% head()
augment(glm_cars) %>% head()
```

```{r}
mtcars_fac <- mtcars %>%
  mutate(cylfac = as.factor(cyl), vs_fac = as.factor(vs))
lm2_cars <- lm(mpg ~ wt + disp + cylfac + vs_fac, data = mtcars_fac)
summary(lm2_cars)

mtcars_fac %>%
  nest_by(cylfac) %>%
  mutate(model = list(lm(mpg ~ wt + disp + vs_fac, data = mtcars_fac))) %>%
  summarise(tidy(model), .groups = "drop")
```

```{r}
library(gapminder)

ggplot(data = gapminder,
       aes(x = year, y = lifeExp, color = continent)) +
  geom_point(position = position_jitter(1,0), size = 0.5) +
  geom_smooth()
```

```{r straight_line}
ggplot(data = gapminder,
       aes(x = year, y = lifeExp, color = continent)) +
  geom_point(position = position_jitter(1,0), size = 0.5) +
  geom_smooth(method = "glm", formula = y ~ x)
```

```{r polynomial}
ggplot(data = gapminder,
       aes(x = year, y = lifeExp, color = continent)) +
  geom_point(position = position_jitter(1,0), size = 0.5) +
  geom_smooth(method = "glm", formula = y ~ poly(x, 2))
```

VOCAB

What might happen if some variables take particular values: counterfactual values.

For example if we know nothing about a new observation, our prediction for that estimate is often based on assuming every variable is at its mean.

Sometimes, we might have very specific questions which require setting (possibly many) combinations of variables to particular values and making an estimate or prediction.

Providing specific estimates, conditional values of covariates, is a nice way to summarize results, particularly for models with unintuitive parameters (i.e. logit models).

```{r sim_data}
ex_dat <- data.frame(num1 = rnorm(200,1,2),
                     fac1 = sample(c(1,2,3), 200, TRUE),
                     num2 = rnorm(200,0,3),
                     fac2 = sample(c(1,2))) %>%
  mutate(yn = num1 * 0.5 + fac1 * 1.1 + num2 * 0.7 +
           fac2 - 1.5 + rnorm(200,0,2)) %>%
  mutate(yb = as.numeric(yn > mean(yn))) %>%
  mutate(fac1 = factor(fac1, labels = c("A", "B", "C")),
         fac2 = factor(fac2, labels = c("Yes", "No")))

```

```{r}
lm_1 <- lm(yn ~ num1 + fac1, data = ex_dat)
lm_1_est <- ggpredict(lm_1, terms = "num1")
lm_1_est
plot(lm_1_est)
```

```{r}
glm(yb ~ num1 + fac1 + num2 + fac2, data = ex_dat, family = binomial(link = "logit")) %>%
  ggpredict(terms = c("num1", 'fac1')) %>% plot()
```

```{r}
glm(yb ~ num1 + fac1 + num2 + fac2, data = ex_dat, family = binomial(link = "logit")) %>%
  ggpredict(terms = c("num1[all]", 'fac1')) %>% plot(facet = TRUE)
```

```{r counterfactual_values}
glm(yb ~ num1 + fac1 + num2 + fac2, data = ex_dat, family = binomial(link = "logit")) %>%
  ggpredict(terms = c("num1[-1,0,1]", 'fac1[A,B]')) %>% plot(facet = TRUE)
```

```{r representative_values}
glm(yb ~ num1 + fac1 + num2 + fac2, data = ex_dat, family = binomial(link = "logit")) %>%
  ggpredict(terms = c("num1[meansd]", 'num2[minmax]')) %>% plot(facet = TRUE)
```

```{r}
lm(yn ~ fac1 + fac2, data = ex_dat) %>%
  ggpredict(terms = c("fac1", "fac2")) %>% plot()
```

```{r}
pander(lm_1)
pander(summary(lm_1))
```

```{r gt_package}
tes_chars <- starwars %>%
  unnest(films) %>%
  unnest(starships, keep_empty = TRUE) %>%
  filter(films == "The Empire Strikes Back") %>%
  select(name, species, starships, mass, height) %>%
  distinct(name, .keep_all = TRUE) %>%
  mutate(starships = ifelse(name == "Obi-Wan Kenobi" | is.na(starships),
                            "No Ship", starships))
glimpse(tes_chars)

tes_chars %>%
  gt()

tes_chars %>%
  group_by(starships) %>%
  gt()

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name")

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name") %>%
  tab_header(
    title = "Star Wars Characters",
    subtitle = "The Empire Strikes Back"
  )

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name") %>%
  tab_header(
    title = "Star Wars Characters",
    subtitle = "The Empire Strikes Back"
  ) %>%
  tab_spanner(
    label = "Vitals",
    columns = vars(mass, height)
  )

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name") %>%
  tab_header(
    title = "Star Wars Characters",
    subtitle = "The Empire Strikes Back"
  ) %>%
  tab_spanner(
    label = "Vitals",
    columns = vars(mass, height)
  ) %>%
  cols_label(
    mass = "Mass (kg)",
    height = "Height (cm)",
    species = "Species"
  )

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name") %>%
  tab_header(
    title = "Star Wars Characters",
    subtitle = "The Empire Strikes Back"
  ) %>%
  tab_spanner(
    label = "Vitals",
    columns = vars(mass, height)
  ) %>%
  cols_label(
    mass = "Mass (kg)",
    height = "Height (cm)",
    species = "Species"
  ) %>%
  fmt_number(
    columns = vars(mass),
    decimals = 0
  )

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name") %>%
  tab_header(
    title = "Star Wars Characters",
    subtitle = "The Empire Strikes Back"
  ) %>%
  tab_spanner(
    label = "Vitals",
    columns = vars(mass, height)
  ) %>%
  cols_label(
    mass = "Mass (kg)",
    height = "Height (cm)",
    species = "Species"
  ) %>%
  cols_align(
    align = "center",
    columns = vars(species, mass, height)
  )

tes_chars %>%
  group_by(starships) %>%
  gt(rowname_col = "name") %>%
  tab_header(
    title = "Star Wars Characters",
    subtitle = "The Empire Strikes Back"
  ) %>%
  tab_spanner(
    label = "Vitals",
    columns = vars(mass, height)
  ) %>%
  cols_label(
    mass = "Mass (kg)",
    height = "Height (cm)",
    species = "Species"
  ) %>%
  cols_align(
    align = "center",
    columns = vars(species, mass, height)
  ) %>%
  row_group_order(
    groups = c("X-wing", "Millennium Falcon")
  )
```

```{r modelsummary}
mod_1 <- lm(mpg ~ wt + disp, data = mtcars)
msummary(mod_1)
```

```{r}
mod_1 <- lm(mpg ~ wt, data = mtcars)
mod_2 <- lm(mpg ~ hp + wt, data = mtcars)
mod_3 <- lm(mpg ~ hp + wt + factor(am), data = mtcars)
model_list <- list("Model 1" = mod_1,
                     "Model 2" = mod_2,
                     "Model 3" = mod_3)
msummary(model_list)
```

```{r}
msummary(model_list, output = "gt") %>%
  tab_header(
    title = "Table 1. Linear Models",
    subtitle = "DV: Miles per Gallon"
  )
```

```{r}
mtcars %>%
  select(1:9) %>%
  tbl_summary()

mtcars %>%
  select(1:9) %>%
  tbl_summary(by = "am")

mtcars %>%
  select(1:9) %>%
  tbl_summary(by = "am") %>%
  as_gt() %>%
  tab_spanner(
    label = "Transmission",
    columns = starts_with("stat_")
  ) %>%
  tab_header(
    title = "Motor Trend Cars",
    subtitle = "Descriptive Statistics"
  )
```

```{r}
corrplot(
  cor(mtcars),
  addCoef.col = "white",
  addCoefasPercent = T,
  type = "upper",
  order = "AOE"
)
```

